Selective electroporation of the viral envelope

### 2021-04-26 documents/lessons_learned.tex
Very large problem size
iLQR

>>>>

I've just realized something.

Experience. When I run a simulation and it errors out, the root cause of that error is *experience*. The plot that I generated and glanced over to inform the next set of parameters is *experience*. The way you pipette 0.5 uL of fluid is experience.

maybe we need to digitize the experience of science.

talk about the electroporation book saying that more effort in standardization is needed

%I hope that software should be written in a manner that is easy to support. This means documentation; reproducible builds;

Not a fan of fortran, but it's actually remarkable in its longevity. A fortran code listing from the 1970s actually has a fair chance of running; has no dependencies, fully self-contained, etc.

### 2021-04-26 documents/pure_electroporation.tex
iLQR

\begin{abstract}
	\noindent The phenomenon of electroporation occurs when extreme applied electric fields in the MV/m range drive transmembrane potentials, leading to pores through membranes; this is commonly harnessed for transfection into the cell in the laboratory, and has recently been shown to be safe and effective in the clinic for ablating tumors.\\

	\noindent As pulse duration is reduced from the microsecond to nanosecond regime, effects targeting smaller intracellular structures are expected and observed, and a degree of tunability arises. Indeed, pulses shorter than 10 ns have been observed to porate the nuclear envelope and organelle membranes while leaving the plasma membrane intact. \\

@@

	\noindent Recently-developed gigawatt pulse systems may offer a possible means of noninvasively propagating such pulses into deep tissues, such as within lungs. 
\end{abstract}

\lettrine{A}{s} an addendum to a technical report \cite{notes2021} by the author presenting experimental data on protein capsid adjacent technique;

In a continued effort to which make ultracrepidarian physicist such as ourselves feel useful indirectly biological

While both mechanisms hinge on time constants, those discussed in the previous paper referred to the viscous loss of energy into the surrounding medium and the normal modes of the viral envelope; here the time constants relate to capacitive charging, and Smoluchowski processes. The relevant damage quantities were mechanical displacement and membrane stress, whereas here we consider transmembrane voltage and pore diameter.

The breakdown of lipid membranes has proved to be an From initial studies of crude membrane breakdown effects \cite{Reversible1979}

"...hit them [cells] with an electric field strong enough to knock over a horse, and they do enough things to justify international meetings, to fill a sizable book, and to lead one to speak of an entirely new technology for cell manipulation."

It is, in general, something of a singular outlier in that it is one of very few biological mechanisms relevant at practical electric field intensities. It also something like integration, which is

@@

At the risk of self-plagarism, we will restate precedent

@@

There is a single report of the electroporation of HIV in the literature. This one of very few reports involving the electroporation of viruses, reporting the ~95\% inactivation of HIV  using a standard commercial Bio-Rad electroporation system at 2.3 MV/m. However, the authors appear to commit such an egregious breach of human research ethics that we cannot cite them here.

@@

Note that, despite clear evidence in [nanosecond review], microsecond in vivo and clinical treatment, nanosecond in vivo conduction, and, The NATO review of vivo data regarding non-contact, "electrodeless", even above the thresholds that might otherwise be expected to cause electroporation. has not yet found any evidence demonstrating electroporation via pulsed RF. de Seze,

@@

In the extreme case of the bacteriophage, the genome is packaged so tightly into the protein capsid that the core is often described as a glassy crystal of DNA\cite{Conformation2007}. ssRNA viruses like coronaviridae are perhaps somewhat less.

On the other hand, the cellular "structome" also contains countless lipid vesicles that might plausibly be damaged; small pockets of lipid in the endoplasmic reticula, Much like in drug discovery, we do not believe we can meaningfully predict all possible off-target effects.

@@

400x difference, reflected in the numerical steady state.

@@

The most obvious difference is scale. Viruses are minute. The classic Schwan equation implies that the transmembrane voltage is proportional to

@@

@@

>>>>

\begin{abstract}
	\noindent The phenomenon of electroporation occurs when extreme applied electric fields in the MV/m range drive transmembrane potentials, leading to pores in, and perhaps the destruction of, membranes; this is commonly harnessed for transfection into the cell in the laboratory, and has recently been shown to be safe and effective in the clinic for the ablation of tumors.

	\noindent As pulse duration is reduced from the microsecond to nanosecond regime, effects targeting smaller intracellular structures are expected and observed, and a degree of tunability arises. Indeed, pulses shorter than 10 ns have been observed to porate the nuclear envelope and organelle membranes while leaving the plasma membrane intact.

@@

	\noindent Recently-developed multi-gigawatt electrical pulse systems may offer a possible means of noninvasively propagating such pulses into deep tissues, such as within lungs. 
\end{abstract}

\lettrine{A}{s} an addendum to a crude technical report \cite{notes2021} by the author, presenting some experimental results on an adjacent technique

In a continued effort to make ultracrepidarian physicists such as ourselves feel useful

While both mechanisms hinge on time constants, those discussed in the previous paper referred to the time for viscous loss of energy into the surrounding medium and the normal modes of the viral envelope; here the time constants relate to capacitive charging, and Smoluchowski processes. The relevant damage quantities were mechanical displacement and membrane stress, whereas here we consider transmembrane voltage and pore diameter.

At the risk of self-plagarism, we will restate some precedent

The field-induced breakdown of lipid membranes has proved to be a particularly useful property of the cell. From initial studies of crude membrane breakdown effects \cite{Reversible1979}

hit "with an electric field strong enough to knock over a horse, and they do enough things to justify international meetings, to fill a sizable book, and to lead one to speak of an entirely new technology for cell manipulation."

It is, in general, something of a singular outlier in that it is one of very few biological mechanisms relevant at practical electric field intensities. Along with thermal treatment, it also performs an operation something like integration, distinct from absolute-value. It is perhaps this property which provides a small degree of tunability.

\footnote{It is amusing to draw parallels between electroporation and Bialke et al's reports on spacecraft reaction wheels. In both cases, destructive process are charging of the external shell (the spacecraft itself), a highly resistive layer (a layer of grease on the ball bearings), and a }

Varactor-like parametric, as the bilayer is squished by the potential;

@@

@@

There is a single report of the electroporation of HIV in the gray literature. This one of very few reports involving the electroporation of viruses, reporting the ~95\% inactivation of HIV using a standard commercial Bio-Rad electroporation system at 2.3 MV/m. However, the authors appear to have committed such an egregious breach of research ethics that we cannot cite them here.

@@

Note that, despite clear evidence in [nanosecond review], microsecond in vivo and clinical treatment, nanosecond in vivo conduction, and, The NATO review concluded that evidence did not support non-contact, "electrodeless", even above the thresholds that might otherwise be expected to cause electroporation. has not yet found any evidence demonstrating electroporation via pulsed RF.

de Seze,

that said, the reviews of the ICNIRP [that 18 GHz  paper]

@@

In the extreme case of the bacteriophage, the genome is packaged so tightly into the protein capsid that the core is often described as a glassy crystal of DNA\cite{Conformation2007}. ssRNA viruses like coronaviridae might be expected to be somewhat more sedate, but this does not seem to be borne out in the core conductivity data.

On the other hand, the cellular "structome" also contains countless vesicles and structures of similar nature that might plausibly be damaged; small pockets of lipid in the endoplasmic reticula - nuclear pore complexes, etc. This appears to be borne out in Gabriel's data; dense tissues exhibit many . It may not be possible to meaningfully predict all possible off-target effects, much like in drug discovery.

@@

The most obvious difference is scale. Viruses are minute. All other properties of the, the classic Schwan equation implies that the steady-state DC transmembrane voltage is proportional to the size of the particle.

Madiyar's Vaccinia electroporation data aligns with. Assuming that the host porates at 0.05 MV/m, the 250 nm vaccinia might be expected to have a threshold of 4 MV/m.

However, this does not explain

%1/(250e-9 / 20e-6)

400x difference, reflected in the numerical steady state simulations.

@@

@@

Differential

@@

two plant
